---
title: "Day 8 - Session 1"
format: html
editor: visual
execute: 
  warning: false
---

## Recap Basics - ggplot

```{r}
library(tidyverse)
library(here)
library(janitor)
library(patchwork)
library(ggrepel)
library(gghighlight)
library(paletteer)
library(ggExtra)
library(ggbeeswarm)
```

read in lizards data

```{r}

lizards <- read_csv(here("data_tidy", "lizards.csv"))
```

## Data Wrangling Practice

Do some data wrangling to create a new subset of lizards where u limit the data to

1.  contain western whiptails, colorado checkered, and eastern fence lizards

2.  Add a new column to convert the length to cm

3.  group. by species and site

4.  return a summary table with mean total length and standard dev of total length by speices and site

```{r}
lizards_subset <- lizards %>% 
  filter(common_name %in% c("western whiptail", "colorado checkered", "eastern fence")) %>%   mutate(total_length_cm = total_length/10) %>% 
  group_by(common_name, site) %>% 
  summarize(mean_length = mean(total_length),
            sd_length = sd(total_length))
```

## Data Vis

make a scatterplot of lizard length vs weight

```{r}
ggplot(lizards, aes(x = total_length, y= weight))+
  geom_point(aes(color = weight), size = 2)+
  scale_y_continuous(limits = c(0,100),
                     breaks = c(seq(from = 10, to = 100, by = 10)),
                     expand = c(0,0))+
  scale_color_gradientn(colors = c("pink","orange","yellow"))
  # scale_color_gradient(low = "pink", high = "yellow")
# +facet_wrap("site") 
# Could add scales = free inside the facet wrap to alter scales
```

Use dplyer::count to find the total number of lizards by common name in the dataset

```{r}
lizards_count <- lizards %>% 
  count(common_name)

ggplot(lizards_count, aes(y = fct_reorder(common_name, n), x = n))+
  geom_col(aes(fill = common_name) )
  
```

fct_reorder: sets the first value as a factor and the second argument gives it the data to decide what order it should be in

## Scales

```{r}

```
